Stack Safety as a Security Property

### 2020-07-05 main.tex
Tweaks

The call stack is a perennially popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
hijacking of the program's control flow. To prevent such attacks, many
software and hardware techniques for stack safety have been proposed, % in the literature and implemented in production systems,
including stack canaries~\citep{Cowan+98},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
split stacks~\citep{Kuznetsov+14},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \apt{Mostly from nick;there could be more}

But what {\em is} this ``stack safety'' of which we speak? Precise formal
definitions are lacking in the literature; instead, the notion of safety
is usually defined negatively, by enumerating the bad things that a safe system
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%with the notion usually being
%defined negatively: through examples of stack-based exploits 
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ... 
We are only aware of one, quite recent, attempt to frame a positive definition.
\citet{Skorstengaard+19} define stack safety as
the conjunction of two properties, local state encapsulation and well-bracketed control flow.
They formalize these properties by defining an abstract capability-based machine   
that ``satisfies [stack safety] by construction.'' We believe there is still room
to craft a more abstract and less intensional characterization of
stack safety, and that is the goal of the present paper. \apt{still needs tuning}

Our approach is to give alternative formal
definitions of local state encapsulation and well-bracketed control flow
using the conceptual framework of language-based security.
%
The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property with both confidentiality and
integrity components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}

% BCP idea: contrast intrinsic vs. extrinsic property definition
% Remember to talk about spatial vs temporal safety

While integrity and confidentiality are standard security
properties, applying them in the stack-safety setting raises a novel challenge:
the notion of ``attacker'' changes during execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be protected from {\em its} callees.
We address this challenge by adjusting the integrity and confidentiality
of different stack frames at each call or return transition between functions.\apt{connection to ``attacker'' still vague.}

The underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety between
functions: each instruction is uniquely owned by some function, and all jumps between functions
are either calls that enter at a marked entry point or returns that re-enter
just after their associated call in the standard nested call structure.
\apt{clarify what we are contributing here.}

To provide evidence that our definitions capture the inituitive notion of stack safety,
we show that they rule out a wide \apt{really?} range of exploitable bad behaviors.\apt{``from the literature''??}
To assess whether they are implementable, 
we compare them against the proposed capability-based enforcement
mechanism of \citet{Skorstengaard+19} \apt{we really should} and the existing tag-based enforcement mechanism
of \citet{DBLP:conf/sp/RoesslerD18}.

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the 
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\Cref{sec:lse-and-wbcf}) in a very general setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\Cref{sec:enforcement}).
  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing property?}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy 
  of \citet{DBLP:conf/sp/RoesslerD18}~(\Cref{sec:lazy}).  This policy does not enforce our property, but we give a
  variant of the policy that does enforce a (provably) weaker variant of our property. \apt{Awkward, but the truth I fear.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with a running example that we refer
to throughout the paper (Section~\ref{sec:running-example}) and some
technical preliminaries (Section~\ref{sec:prelim}).

\section{Running Example}
\label{sec:running-example}

@@

@@

@@

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to 18}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 6]}{r_5 \leftarrow 0}
      {Load argument {\tt z} $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}      
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as ther return of {\tt g}}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{8}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace
  \rb{A suggestion for the description of $\jal$ and $\jalr$:
    call(function label) [if labels are added next to the entrypoints]
    or call(address of entrypoint) and return()?}
}
\label{fig:running-trace-b}
\end{figure}

@@

\leo{TODO: Add citations of mitigation techniques}

@@

\paragraph*{Reusing call frames}

\leo{TODO}

\section{Machines, Traces, and Observations}
\label{sec:prelim}

In this section we describe our machine model. Our definition of stack
safety is generic and this section discusses an abstract interface to
our machine model and some kind of static or dynamic, software- or
hardware-enforced security monitor. We will refer back to the running
example of the previous section for concreteness, and then,
in~\Cref{sec:micropolicies}, extend this model with a dynamic policy
enforcement mechanism in~\Cref{sec:micropolicies} based on
\bcp{NickPaper,AndMaybeSomeOthers}.

\subsection{Values, States, and Observations}

@@

The step function of the machine takes a machine state and produces a
pair of a machine state and an {\em observation}, which may be either
silent (written $\tau$) or a word of data that the machine
communicates to the outside world (e.g., by writing to a DMA
address):
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
%

Mapping a conventional architecture to this abstract interface is
straightforward. For the running example of the previous section,
$\WORDS$ is the set of 64-bit integers, $\REGS$ is the set $\{r_0,
\cdots, r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$. Observations are
the writes to the special $\rout$ location: if during an execution
step a word $w$ is written to $\rout$ then $\obs = w$, otherwise $\obs
= \tau$.

\leo{Other choices for observations? Where should that be discussed?}
\bcp{But in this paragraph we are describing the particular
choise that we make for our running example, no?  But we certainly want to
discuss this fully someplace!}

\subsection{Policies}

@@

\leo{Policies of interest: null, micro, cherri, software-only}
\leo{TODO: Write down carefully}

\subsection{Traces}

\leo{Which of the machine, machine-observation, machine-policy,
  and machine-policy-observation traces should we distinguish
  in metavariables?}

A trace over a type $A$ is a nonempty, finite or infinite sequence of
$A$s. We will mostly be concerned with sequences of machines states,
denoted $\machT$, and with sequences of triples of (machine states,
policy states, and observations), usually denoted $\MPT$.  \rb{The way
  the definitions are set up, one may expect MPO, but\ldots} We
will require four generic operations on traces.

\bcp{TODO: Let's also try to clean up the spacing a little -- many of these
  sequences of rules can be typeset on fewer lines}

\paragraph*{Trace-Of}

The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:  
%
\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
%
\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

In our running example in Figure~\Cref{fig:running-example-b} we can
see a depiction of such a trace: starting from an initial machine
state at the very top, each instruction constructs a new machine state
(the diff of which compared to the previous one appears in the
``Effect'' column) and the majority of observations are $\tau$ with
the exception of instructions 31 and 13 which are writes to $\rout$
and where the observations are $0$ and $60$ respectively.

In a similar fashion, we define \(\hookrightarrow_P\) to relate a
machine-policy state pair to the trace of triples of machine states,
policy states, and observations induced by by \(\stepstoobs{}_P\).
Unlike the $\MOTRACE$ above, this trace can be finite as the
policy can cause a fault:
%
\[(m,p) \hookrightarrow_P MP : \MACHS \times \POLS \rightarrow \MPOTRACE \]
%
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
%
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
  all these.  E.g., do these traces include the final state of a finite
  execution?  If so, what observation is it paired with?}\leo{The final
  observation is a $\tau$ based on the rule in the inlined figure, so
  it is essentially ignored. Point that out somehow?}

\paragraph*{Prefix-Up-To}
The ``prefix-up-to'' operation \(\PUT(\Trace,f)\) \bcp{Wouldn't
  ``While'' be more mnemonic?}  takes a trace \(T\) and a predicate on
elements \(f\) and gives the prefix of \(T\) ending with the first
element on which \(f\) holds, coinductively:

\begin{center}
\begin{minipage}{.3\textwidth}
\judgment{\(f\ \traceelem\)}
         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
\end{minipage}
%
\begin{minipage}{.3\textwidth}
\judgment{}
         {\(\PUT(f,\traceelem) = \traceelem\)}
         %
\end{minipage}                  
\begin{minipage}{.3\textwidth}
\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
\end{minipage}
\end{center}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds,
\(\PUT\) is the identity function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
  think.}  \leo{There were inconsistencies which i fixed. I wanted the
  metavariables defined to be consistent and the ``temporary''
  variable which is the full trace to be primed. Should I swap those?}

@@

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT'\)}
            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT') = \MPT\)}
         {\(\mpstatename \hookrightarrow \MPT | f\)}

@@

In this section we will describe stack safety as a formal property. We
begin with our model of \leo{compiler-provided?} knowledge about the
program structure, and then dive in to formal definitions local state
encapsulation and well-bracketed control flow.

@@

>>>>

The call stack is an increasingly\apt{``a perennially'' ?} popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from attackers reading or corrupting private stack data, to
highjacking a program's control flow. To prevent such attacks, many
techniques have been proposed in the literature\leo{cite, point out?
  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
  stronger point.} aiming to achieve ``stack
  safety''.

But what {\em is} this stack safety of which we speak? A precise formal
definition is lacking in the literature, with the notion usually being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other
  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\apt{is this gone?}
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new
  part?}\bcp{If it is not too complicated to explain, we should definitely
  put it
  here.  If it is complicated, we can just say in text that there are other
  sorts of attacks too, and that we deal with them, and give a forward
  pointer.}
Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
  we are only aware of one, quite recent, attempt to do something better} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
but even here, stack safety is defined with respect to an abstract
machine that ``satisfies stack safety by construction,'' begging
the question of what {\em is} the property that holds by construction?
\apt{Suggest to move first sentence of next para up to this one.}

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. The first
contribution of the present paper is to
turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
  claim this} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully
developed.}\bcp{Yes, and let's remember that they are highly likely to be
reviewing this paper!  Let's be as generous as we can be without
under-selling our delta.}

The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property with both confidentiality and
integrity components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}

While integrity and confidentiality are relatively standard security
properties, the stack-safety setting raises an interesting additional
challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be proteced from {\em its} callees.
We address this challenge by varying the observation power\apt{is this
  the right word? doesn't it conflict with our later use of ``observation''?}
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
  the sentence are about two very different things.  The second deserves its
  own paragraph, IMO.}\apt{+1}

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). Well-bracketed control flow establishes safety between
functions: all code is uniquely owned by some function, and all jumps between functions
are either calls that enter at a marked entry point or returns that enter
just after their associated call in the standard nested call structure.

%\leo{There needs to be some informal description
%  of the wbcf part here. Sean?}

In summary, we offer the following contributions\apt{is this up to date with current paper organization?}:
\begin{itemize}
\item \leo{More? Extensional}\bcp{We might need to invent a word for this---''quasi-extensional''?} formal characterizations of the
  intuitive concepts of local state encapsulation and
  well-bracketed control flow~(\Cref{sec:lse-and-wbcf}).  As evidence that this
  property is strong enough to capture the folk intuition, we show
  that the stack unsafe examples from the literature also fail our
  property~(\Cref{sec:disc}).\apt{Something about relation to StkTokens formalization?}
\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
  realizible by
  a previously proposed micropolicy enforcement
  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
    substantial body of code}. We also present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker (\Cref{sec:lazy}).
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework. \bcp{Which section?}
\end{itemize}
Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
%
We begin with a running example that we're going to keep coming back
to throughout the paper (Section~\ref{sec:running-example}) and some
technical preliminaries (Section~\ref{sec:prelim}).

\section{Stack Safety and Unsafety by Example}
\label{sec:running-example}

@@

\subsection{Running Example}

@@

\bcp{Notation nit: in many assembly notations, square brackets mean
  ``{\em indirect} through this location...''  And a very small nit: Change
  $\rsp-2$ to $\rsp\mathord{-}2$}

@@

\setcounter{pcctr}{27}
% H - Body
\vspace*{0.2em}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ r_4 ~ \rsp ~ 0}{r_4 \leftarrow [\rsp]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[105] \leftarrow 18}
      {Set {\tt w} to 18}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 106}
      {Deallocate {\tt h}'s call frame}
  \row{\lw ~ r_5 ~ \rsp ~ \negate 6}{r_5 \leftarrow [\rsp - 6]}{r_5 \leftarrow 0}
      {Load argument {\tt z} $r_5$}
  \row{\sw ~ \rout ~ r_5 ~ 0}{[\rout] \leftarrow r_5}{[99] \leftarrow 0}
      {Print {\tt z}}      
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 18}
      {Add $r_4$ and $r_5$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[103] \leftarrow 18}
      {Store $r_4$ as ther return of {\tt g}}
  \row{\lw ~ \rra ~ \rsp ~ \negate 2}{\rra \leftarrow [\rsp - 2]}{\rra \leftarrow 8}
      {Load return address}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 104}
      {Deallocate {\tt g}'s local state}
  \row{\jalr ~ \rra ~ \rra ~ 0}{??}{PC \leftarrow 8}{Return to {\tt main}}
\end{array}
\]
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{43.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{2}{green}[{\makebox[0pt]{{\tt g}'s call frame }}]%
\memory{6}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{0}
\MemoryLabel{-47em}{0.75em}{42}
\MemoryLabel{-43em}{0.75em}{0}
\MemoryLabel{-39em}{0.75em}{0}
\MemoryLabel{-35em}{0.75em}{18}
\\
\end{center}
% Register File
\summary{8}{0}{\thepcctr}{104}{18}{0}
%\vspace*{0.2em}
\setcounter{pcctr}{8}
\[
\begin{array}{c|c|c|c|c}
  \row{\lw ~ \rsp ~ r_4 ~ \negate 1}{r_4 \leftarrow [\rsp - 1]}{r_4 \leftarrow 18}
      {Load return value to $r_4$}
  \row{\sw ~ \rsp ~ r_4 ~ \negate 3}{[\rsp - 3] \leftarrow r_4}{[101] \leftarrow 18}
      {Set {\tt y} to the result}
  \row{\addi ~ \rsp ~ \rsp ~ \negate 4}{\rsp \leftarrow \rsp - 2}{\rsp \leftarrow 102}
      {Deallocate {\tt g}'s call frame}
  \row{\lw ~ \rsp ~ r_5 ~ \negate 2}{r_5 \leftarrow [\rsp - 2]}{r_5 \leftarrow 42}
      {Load {\tt x} to $r_5$}
  \row{\add ~ r_4 ~ r_4 ~ r_5}{r_4 \leftarrow r_4 + r_5}{r_4 \leftarrow 60}
      {Add {\tt x} and {\tt y}}
  \row{\sw ~ \rout ~ r_4 ~ 0}{[\rout] \leftarrow r_4}{[99] \leftarrow 60}
      {Print {\tt x + y}}
\end{array}
\]
\vspace*{0.2em}
% Return to main
% Memory Layout
\begin{center}
\MemoryLabel{4em}{2em}{0}
\MemoryLabel{21em}{2em}{99}
\MemoryLabel{31.5em}{2em}{\SP}
\memory{3}{gray}[{\makebox[0pt]{Instructions}}]%
\hspace*{3pt}
$\cdots$
\memory{1}{white}[{\makebox[0pt]{Output}}]%
\memory{2}{green}[{\makebox[0pt]{${\texttt main}_{priv}$}}]%
\memory{8}{gray}%[{\makebox[0pt]{Uninitialized}}]
~$\cdots$
\MemoryLabel{-51em}{0.75em}{60}
\MemoryLabel{-47em}{0.75em}{60}
\MemoryLabel{-43em}{0.75em}{42}
\\
\end{center}
% Starting Register File
\summary{\thepcctr}{0}{37}{102}{60}{42}
\caption{Detailed Execution Trace}
\label{fig:running-trace-b}
\end{figure}

@@

@@

\section{Background}
\label{sec:prelim}

To set the stage, we first describe our machine model. Our definition
of stack safety is generic and this section discusses an abstract
interface to our machine model and some kind of static or dynamic,
software- or hardware-enforced security monitor.  For concreteness,
our running examples will use a conventional machine architecture and
a standard (RISC-V) instruction set; in~\Cref{sec:micropolicies} we
will extend this model it with a dynamic policy enforcement mechanism
based on \bcp{NickPaper,AndMaybeSomeOthers}.

\paragraph*{Values, States, and Observations}

@@

The step function of the
machine takes a machine state and produces a pair of a
machine state and an {\em observation}, which may be either
silent (written $\tau$) or a word of data that the machine communicates to
the outside world (e.g., by writing to a DMA address):\leo{Needs more
  exposition}\bcp{Like that?}
%
\[\obs \in \OBSS = \{\tau\} \uplus \WORDS\]
\[\mach \stepstoobs{\obs} \mach' : \MACHS \rightarrow \MACHS \times \OBSS \]
%

\paragraph*{Running Example: RISC-V}

Mapping a conventional architecture to this abstract interface is
straightforward. To build up our running example, we are going to use
a standard 64-bit RISC machine, just like the one we are using in our
experiments later on \leo{Mention FORVIS?}. In this setting, $\WORDS$
is the set of 64-bit integers, $\REGS$ is the set $\{r_0, \cdots,
r_{31}\} \cup \{\PCname\}$, and $\SP = r_2$.
% \leo{Mention We use the standard encoding of instructions to words?}
% BCP: Nah.

The first interesting choice is that of observations, as there are
multiple reasonable options. At one extreme, we could choose to define
observations as the value written at any component. However, that
would pose problems later on when we treat some components as ``secret''
when they correspond to a caller's private state. \leo{On the other extreme,
we could pick a particular component and treat it as the ``output'' of
our system. }  \bcp{But in this paragraph we are describing the particular
choise that we make for our running example, no?  But we certainly want to
discuss this fully someplace!}

\paragraph*{Policies}

@@

\paragraph{Traces}

A trace over a type $A$ is a nonempty, finite or infinite
sequence of $A$s. We will mostly be concerned with sequences of
machines states, usually\bcp{not always?} denoted $\machT$, and with sequences of
triples of (machine states, policy states, and observations), usually
denoted $\MPT$.
\rb{The way the definitions are set up, one may expect MPO, but\ldots}
To continue\bcp{??}, we will require four generic operations
on traces: a ``trace-of'' operation to construct the
trace induced by a step relation and a machine state
\rb{not an op on traces per se?}\bcp{+1};
a ``split'', to break a
trace into two components based on a predicate; a ``join'', the
reverse of split; and a ``prefix'', as a shorthand to only keep the
first component of a split.  \bcp{Not sure it's worth enumerating them
  rather than just getting down to the actual definitions below.}

The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:  \bcp{above, we
  showed operators applied to metavariables, not just the bare operator
  symbol.}
%
\[\hookrightarrow : \machT \rightarrow \MOTRACE\]%
%
\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

\leo{TODO: Add an example of a trace once I figure out the notations}
\bcp{Let's also try to clean up the spacing a little -- many of these
  sequences of rules can be typeset on fewer lines}

Similarly, we define \(\hookrightarrow_P\) to relate a machine-policy
state pair to the trace of triples of machine states, policy states,
and observations induced by by \(\stepstoobs{}_P\): \bcp{Observe that this
  time the trace can be either finite or infinite.}
%
\[\hookrightarrow_P : \MACHS \times \POLS \rightarrow \MPOTRACE \]
%
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
%
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
  all these.  E.g., do these traces include the final state of a finite
  execution?  If so, what observation is it paired with?}

% SNA: fixed
%\leo{PUT is completely ambiguous as a choice of abbreviation. Just ``Pre''
%  seems better}
The {\it prefUpTo} operation \(\PUT(\Trace,f)\) \bcp{Why does it have two
  different abbreviated names?  And wouldn't ``While'' be more mnemonic?}
takes a trace \(T\)
and a predicate on elements \(f\) and gives the prefix of \(T\)
ending with the first element on which \(f\) holds, coinductively:

\judgment{\(f\ \traceelem\)}
         {\(\PUT(f,\traceelem\Trace) = \traceelem\)}
%
\judgment{}
         {\(\PUT(f,\traceelem) = \traceelem\)}
%
\judgmenttwo{\(\neg f\ \traceelem\)}{\(\PUT(f,\Trace) = \Trace'\)}
            {\(\PUT(f,\traceelem\Trace) = \traceelem\Trace'\)}
%
Note that if the resulting trace is a strict prefix of the parameter, then the
predicate must hold on its final element. If the predicate never holds,
\(\PUT\) is the identity function.

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT'\) is the prefix of
running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I think.}

@@

\judgmenttwo{\(\mpstatename \hookrightarrow \MPT\)}
            {\(\PUT(\lambda \mpostate{\mach}{\pol}{\obs}.f(\mach),\MPT) = \MPT'\)}
         {\(\mpstatename \hookrightarrow \MPT' | f\)}

@@

In this section we will describe stack safety as a formal property. We
begin with introducing a running example that will help keep things
concrete throughout this section, continue with our model of
\leo{compiler-provided?} knowledge about the program structure, and
then dive in to formal definitions local state encapsulation and
well-bracketed control flow.

@@

\section{Discussion}
\bcp{Make this a subsection?}
\label{sec:disc}

  \paragraph{Examples}
